\documentclass[11pt, letterpaper]{article}

% --- Packages ---
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{caption}
\usepackage{float}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning}
\usepackage{pgfplots} % For native curve fitting and graphing
\pgfplotsset{compat=1.18}
\usepackage{hyperref}

% --- Title Section ---
\title{\textbf{Investigation of Pulse Propagation in Transmission Lines and Coaxial Cables}}
\author{Akhilesh Menon, Adam Ben Lassoued, Fola Adeleke}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
This experiment investigated the finite propagation speed of electrical signals and impedance matching using a 41-unit LC discrete transmission line model and continuous coaxial cables. By analyzing the delay time of transmitted and reflected pulses, the propagation speed in the LC network was measured at $(2.55 \pm 0.01) \times 10^5 \text{ units/s}$, closely matching the theoretical expectation of $2.58 \times 10^5\ \text{units/s}$. A \qty{48.36}{m} coaxial cable yielded a signal velocity of $1.96 \times 10^8 \text{ m/s}$. The characteristic impedance of the coaxial cable was experimentally determined to be $Z_0 \approx 51\ \Omega$. Furthermore, signal attenuation was calculated to be $-0.026 \text{ dB/m}$ for the \qty{48.36}{m} cable.
\end{abstract}

\section{Introduction}
Standard circuit theory simplifies networks by assuming the speed of propagation of information is infinite. However, high-frequency changes do not instantaneously react across a circuit. In reality, pulses travel at speeds constrained by the physical properties of the medium.

Transmission lines can be modeled as discrete ladder networks of inductors ($L$) and capacitors ($C$). Analyzing the rates of change for voltage and current yields the wave equations:
\begin{equation}
\frac{\partial^{2}V}{\partial x^{2}}=L_{0}C_{0}\frac{\partial^{2}V}{\partial t^{2}} \quad \text{and} \quad \frac{\partial^{2}I}{\partial x^{2}}=L_{0}C_{0}\frac{\partial^{2}I}{\partial t^{2}}
\end{equation}
From these, the theoretical velocity of wave propagation is $v = 1/\sqrt{L_0C_0}$. A critical property of these lines is the characteristic impedance ($Z_0$), defined as the ratio of voltage to current ($Z_0 = \sqrt{L_0/C_0}$). When a transmission line is terminated by a load $Z_L$, a reflection occurs if $Z_L \neq Z_0$. The reflection coefficient is given by $r = \frac{Z_L - Z_0}{Z_L + Z_0}$. This experiment verifies these properties using a physical LC model and standard coaxial cables.

\section{Experimental Setup and Technique}
The apparatus consisted of a KEYSIGHT EDU33211A waveform generator, a digital oscilloscope, an LC transmission board (41 units, $L = \qty{1.5}{mH}$, $C = \qty{0.01}{\mu F}$), and several spools of coaxial cable.

A TEE connector split the pulse generator output; one path connected directly to Channel 1 of the oscilloscope to monitor the input pulse, and the other path fed into the transmission line. The far end of the transmission line was connected to Channel 2 to observe transmission delay or left open/shorted to observe reflections on Channel 1.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    box/.style={draw, rectangle, minimum width=2.2cm, minimum height=0.9cm,
                align=center, font=\small},
    arr/.style={-{Stealth}, thick},
    node distance=1.5cm and 2.0cm
]
% Nodes
\node[box] (gen)  {Pulse\\Generator};
\node[box, right=of gen] (tee)  {TEE\\Connector};
\node[box, right=of tee] (line) {Transmission\\Line / Cable};
\node[box, right=of line] (term) {Termination\\($Z_L$)};
\node[box, above=of tee] (ch1)  {Oscilloscope\\Ch.\,1};
\node[box, above=of term] (ch2)  {Oscilloscope\\Ch.\,2};

% Arrows
\draw[arr] (gen)  -- node[above, font=\footnotesize]{BNC} (tee);
\draw[arr] (tee)  -- node[above, font=\footnotesize]{coax} (line);
\draw[arr] (line) -- (term);
\draw[arr] (tee)  -- node[right, font=\footnotesize]{monitor} (ch1);
\draw[arr] (term) -- node[right, font=\footnotesize]{Ch.\,2 / open} (ch2);

% Reflection arrow — routed below the diagram as a clean right-angle path
\draw[arr, dashed]
    (term.south) -- ++(0,-0.6) -- node[below, font=\footnotesize]{reflected pulse}
    ++(-6.0,0) -- (tee.south);
\end{tikzpicture}
\caption{Block diagram of the experimental setup. The TEE connector splits the generator output so that the incident pulse is monitored on Ch.\,1 while the transmitted (or reflected) pulse is observed on Ch.\,2.}
\end{figure}

\section{Data and Results}

\subsection{Raw Data}
Table 1 outlines the delay time ($\Delta t$) across varying nodes of the LC line. Table 2 details the signal behavior for the \qty{48.36}{m} coaxial cable under different resistive terminations, and is used for all subsequent coaxial calculations.

\begin{table}[H]
\centering
\caption{Delay time vs. Number of LC Units ($\pm 1 \mu s$)}
\begin{tabular}{cc|cc|cc}
\toprule
$N$ (Units) & Delay ($\mu s$) & $N$ (Units) & Delay ($\mu s$) & $N$ (Units) & Delay ($\mu s$) \\ \midrule
41 & 160 & 33 & 130 & 25 & 99 \\
40 & 156 & 32 & 126 & 24 & 94 \\
39 & 153 & 31 & 121 & 23 & 90 \\
38 & 150 & 30 & 119 & 22 & 85 \\
37 & 145 & 29 & 113 & 21 & 81 \\
36 & 142 & 28 & 110 & 20 & 77 \\
35 & 138 & 27 & 107 & 19 & 74 \\
34 & 134 & 26 & 103 &    &    \\ \bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Coaxial Cable Terminations (Length: $48.36 \pm 0.02$ m)}
\begin{tabular}{lcccc}
\toprule
Termination & Delay ($\pm 2$ ns) & $V_{signal}$ ($\pm 0.05$ V) & $V_{return}$ ($\pm 0.05$ V) & Exp. $r$ ($V_{ret}/V_{sig}$)\\ \midrule
Open Circuit & 494 & 4.19 & 3.12 & 0.74 \\
Closed Circuit & 494 & 4.19 & -3.11 & -0.74 \\
$270\ \Omega$ & 494 & 4.21 & 2.13 & 0.51 \\
$100\ \Omega$ & 494 & 4.21 & 0.97 & 0.23 \\
$51\ \Omega$ & 494 & 4.21 & -0.11 & -0.03 \\
$27\ \Omega$ & 494 & 4.21 & -2.64 & -0.63 \\ \bottomrule
\end{tabular}
\end{table}

\subsection{Uncertainty of Direct Measurements}
Direct measurements carried inherent instrumental uncertainties. Time delays measured via the oscilloscope cursors had an estimated uncertainty of $\pm \qty{1}{\mu s}$ for the LC line and $\pm \qty{2}{ns}$ for the coaxial cables due to sampling rate limits. Voltage amplitudes carried a $\pm \qty{0.05}{V}$ uncertainty based on the vertical scale resolution. Coaxial cable lengths were taken as marked, with an estimated $\pm \qty{0.02}{m}$ uncertainty due to the physical connectors.

\section{Data Analysis and Calculations}

\subsection{Curve Fitting and LC Speed of Propagation}
The measured delay for the LC units represents the transit time. By plotting delay $\Delta t$ against the number of units $N$, the slope $m$ yields $1/v$. 

\begin{figure}[H]
\centering
\begin{tikzpicture}
\begin{axis}[
    title={Delay Time vs. Number of LC Units},
    xlabel={Number of LC Units ($N$)},
    ylabel={Delay Time ($\mu s$)},
    grid=major,
    width=0.8\textwidth,
    height=0.4\textwidth,
    legend pos=north west,
    error bars/y dir=both,
    error bars/y explicit
]
\addplot[only marks, color=blue, mark=*, error bars/.cd, y dir=both, y fixed=1] coordinates {
    (19,74) (20,77) (21,81) (22,85) (23,90) (24,94) (25,99) (26,103) (27,107) (28,110) 
    (29,113) (30,119) (31,121) (32,126) (33,130) (34,134) (35,138) (36,142) (37,145) 
    (38,150) (39,153) (40,156) (41,160)
};
% Line of best fit: slope = 3.92, intercept = 0 (forced through origin)
\addplot[thick, red, domain=18:42] {3.92*x};
\legend{Experimental Data, Linear Fit ($t = 3.92\, N$)}
\end{axis}
\end{tikzpicture}
\caption{Linear regression of signal delay through discrete LC units. The error bars ($\pm \qty{1}{\mu s}$) are plotted but are smaller than the data markers.}
\end{figure}

\textbf{Goodness of Fit:} The linear regression (forced through the origin, as $t = 0$ when $N = 0$) yielded a highly correlated fit with an $R^2 \approx 0.999$ and a Root Mean Square Error (RMSE) of $\approx \qty{0.8}{\mu s}$, which falls within our direct measurement uncertainty.

The slope and its standard error from the regression are $m = (3.92 \pm 0.01) \times 10^{-6}\ \text{s/unit}$. Therefore, the experimental speed is:
\begin{equation}
v_\text{exp} = \frac{1}{m} = \frac{1}{3.92 \times 10^{-6}} = 2.55 \times 10^5 \text{ units/s}
\end{equation}
The uncertainty in $v_\text{exp}$ is propagated via $\delta v = \delta m / m^2$:
\begin{equation}
\delta v_\text{exp} = \frac{\delta m}{m^2} = \frac{0.01 \times 10^{-6}}{{\left(3.92 \times 10^{-6}\right)}^2} \approx 0.65 \times 10^3\ \text{units/s} \approx 0.01 \times 10^5\ \text{units/s}
\end{equation}
Thus the experimental speed is $v_\text{exp} = (2.55 \pm 0.01) \times 10^5\ \text{units/s}$.
Theoretical speed: $v_{theo} = 1/\sqrt{LC} = 1/\sqrt{(1.5\ \text{mH})(0.01\ \mu\text{F})} = 2.58 \times 10^5\ \text{units/s}$. The experimental value is within $1.2\%$ of the theory.

\subsection{Coaxial Cable Propagation and Attenuation}
For the coaxial cables, the wave travels round-trip during reflection measurements ($d = 2L$). Using the \qty{48.36}{m} open-circuit data from Table~2 ($t = \qty{494}{ns}$, $V_{sig} = \qty{4.19}{V}$, $V_{ret} = \qty{3.12}{V}$):
\begin{equation}
v_{coax} = \frac{2L}{t} = \frac{2 \times 48.36}{494 \times 10^{-9}} = 1.96 \times 10^8 \text{ m/s}
\end{equation}
This agrees with the theoretical speed in a dielectric ($\epsilon_r = 2.25$):
$v_{theo} = c/\sqrt{\epsilon_r} = 3\times 10^8 / 1.5 = 2.00 \times 10^8 \text{ m/s}$.

Attenuation for the \qty{48.36}{m} cable (round trip) was calculated using:
\begin{equation}
\text{Attenuation} = 20 \log_{10}\!\left(\frac{V_{ret}}{V_{sig}}\right) = 20 \log_{10}\!\left(\frac{3.12}{4.19}\right) = -2.56 \text{ dB}
\end{equation}
Dividing by the total travel distance ($2 \times 48.36 = \qty{96.72}{m}$) yields a specific attenuation of $\approx -0.026 \text{ dB/m}$.

\subsection{Propagation of Uncertainty}
The uncertainty in the theoretical LC speed is derived using partial derivatives of $v = {(LC)}^{-1/2}$:
\begin{equation}
\delta v_{theo} = v_{theo} \sqrt{ {\left( -\frac{1}{2} \frac{\delta L}{L} \right)}^2 + {\left( -\frac{1}{2} \frac{\delta C}{C} \right)}^2 }
\end{equation}
Assuming standard component tolerances of $5\%$ for $L$ and $10\%$ for $C$:
\begin{equation}
\frac{\delta v}{v} = \sqrt{ {(-0.5 \times 0.05)}^2 + {(-0.5 \times 0.10)}^2 } = \sqrt{0.000625 + 0.0025} \approx 5.6\%
\end{equation}
Thus, theoretical speed is $(2.58 \pm 0.14) \times 10^5$ units/s. The experimental speed falls well within this range.

\section{Sources of Uncertainties}
1. \textbf{Instrumental Limits:} The dominant uncertainty arises from the digital quantization of the oscilloscope's analog-to-digital converter, limiting precision in determining the exact start of a pulse.
2. \textbf{Non-Ideal Components:} The LC mathematical model assumes zero electric resistance. In reality, inductors possess inherent Equivalent Series Resistance (ESR), causing minor signal dispersion and contributing to deviations from ideal behavior.
3. \textbf{Connectors:} The physical length of the BNC TEE connectors and internal wiring adds parasitic capacitance and marginal length not accounted for in the printed spool lengths.

\section{Discussion and Conclusion}
The results strongly confirm the theoretical models for signal propagation. Impedance matching was vividly demonstrated; as seen in Table 2, inserting a $51\ \Omega$ termination reduced the reflected pulse to nearly zero ($r \approx -0.03$), perfectly aligning with the standard $Z_0 \approx 50\ \Omega$ specification for RG-58 cables. 

When the cable was open-circuited ($Z_L = \infty$), the reflection was non-inverted ($r>0$), and when short-circuited ($Z_L = 0$), it was perfectly inverted ($r<0$). We safely ignored the oscilloscope's presence during reflection measurements because its input impedance ($1\ M\Omega$) acts virtually as an open circuit in parallel with the $50\ \Omega$ line, drawing negligible current. Furthermore, considering the DC equivalent of an open-circuited line, the potential difference across the circuit will simply equal the source voltage, as no current flows and therefore no voltage is dropped across the line's internal resistance.

Ultimately, pulses do not travel at $c$, but are slowed by the dielectric constant of the insulator. The experiment successfully mapped these propagation speeds, characterized $Z_0$, and quantified real-world signal attenuation.

\end{document}